---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First I will load the data into R:

```{r}
rawData <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

I will now make a collapsed dataset that reports the mean and median by day:

```{r, echo=FALSE}
meanByDay<-aggregate(rawData$steps, by = list(rawData$date), FUN = mean, na.rm = T)
medianByDay<-aggregate(rawData$steps, by = list(rawData$date), FUN = median, na.rm = T)

cbind(meanByDay, medianByDay[,2])
```

I will now generate a histogram showing the total number of steps taken per day

```{r}
byDay <- aggregate(rawData$steps, by=list(rawData$date), FUN = sum, na.rm = T)
library(ggplot2)
qplot(byDay$x, geom='histogram', data=byDay,  binwidth = 1000, xlab='Number of steps')

```

## What is the average daily activity pattern?

Below is a time-series plot by 5 minute interval throughout the day, averaged over each day:

```{r}
z <- aggregate(rawData$steps, by = list(rawData$interval), FUN = mean, na.rm=T)

max(z$x)


g <- ggplot(z, aes(Group.1, x)) + geom_line() + xlab('5-minute time intervals')
g
```

## Imputing missing values

The total number of missing values in the dataset is:

```{r}
sum(is.na(rawData))
```

Now I will replace all NA's with the mean for that day.  I will use this new dataset to calculate a new mean for each day as well as the corresponding histogram of the sum. 

```{r}


temp<-merge(rawData, meanByDay, by.x='date', by.y='Group.1')

for(i in 1:length(temp[,1])){
  
  if(is.na(temp[i,2])){
    temp[i,2] <- temp[i,4]
  }
}

aggregate(temp$steps, by = list(temp$date), FUN = mean)

m <- aggregate(temp$steps, by = list(temp$date), FUN = sum)

ggplot(m, aes(x)) + geom_histogram(binwidth=1000)

```


## Are there differences in activity patterns between weekdays and weekends?


Finally I will split the data by weekend or weekday and mike a final plot:

```{r}

rawData$weekday <- weekdays(as.Date(rawData$date))
rawData$recoded <- ifelse(rawData$weekday == "Saturday" | rawData$weekday == "Sunday",'Weekend','Weekday')

g <- aggregate(rawData$steps, by = list(rawData$interval, rawData$recoded), FUN = mean, na.rm=T)

ggplot(g, aes(Group.1, x)) + geom_line() + facet_grid(Group.2 ~ .) + xlab('5-minute intervals')
```















